Towards a miniKanren with fair search strategies

### 2019-05-26 Paper/paper.tex
Just a few loose ends

\NOTE[DPF]{``current'' was for BFSser}
\NOTE[LKC]{Do you mean I should not reserve ``current'' for BFSser, or I should 
do so?}

@@

The miniKanren in TRS2 implements interleaving DFS (\DFSi), the cause of this 
unexpected result. With this search strategy, each \conde{} clause takes half 
of its received computational resources and passes the other half to its 
following clauses, except for the last clause that takes all resources it 
receives. In the example above, the \texttt{a} clause takes half of all 
resourses. And the \texttt{b} clause takes a quarter. Thus \texttt{c} and 
\texttt{d} barely contribute to the result.

@@

Given the following program, it is natural to expect lists of each
letter to constitute $1/4$ in the answer list. \DFSi, TRS2's search
strategy, however, results in many more lists of \texttt{a}s than
lists of other letters. And some letters (e.g. \texttt{c} and
\texttt{d}) are rarely seen. The more clauses the worse the situation.

@@

This implementation can be improved in two ways. First, as mentioned in 
\autoref{sec:fairconj}, \BFS{} puts answers in buckets and answers of the same 
cost are in the same bucket. In the above implementation, however, 
it is not obvious how we manage cost information---the \texttt{car}s of a 
space have cost 0 (i.e., they are all in the same bucket), and every 
thunk indicates an increment in cost. It is even more subtle that 
\appendInfFair{} and the \appendMapInfFair{} respects the cost information. 
Second, \appendInfFair{} is extravagant in memory usage. It makes $O(n+m)$ new 
\texttt{cons} cells every time, where $n$ and $m$ are the sizes of 
the first buckets of two input spaces. \DFSf{} also has the space extravagance.

@@

\NOTE[LKC]{I delete the ending sentence of the above paragraph.}

@@

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since the result of \texttt{(run n ? g)} is 
the same either way. (See supplements for the formal proof.) To 
compare efficiency, we translate \BFSser{}'s Haskell code into Racket (See 
supplements for the translated code). The translation is direct due to the 
similarity of the two relational programming systems. The translated code is 
longer than \BFSimp{}. And it runs slower in all benchmarks. Details about 
differences in efficiency are in \autoref{sec:quan}.

\NOTE[DPF]{Make another macro or use a number when we want to drop ``the''
from ``the improved BFS''.  In the table below, use that macro.}
\NOTE[LKC]{I have defined another macro BFSim, which doesn't start with 
``the''. Also, I have corrected the table below.}

@@

\NOTE[DPF]{tilde problem 7 lines down.}
\NOTE[LKC]{Corrected.}

@@

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}, the current standard miniKanren 
search strategy. 
Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
or better. 
It is hard to conclude how \disj{} fairness affects performance based on these 
statistics.
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

\NOTE[DPF]{Please leave ``prove'' and don't use ``proof''.}
\NOTE[LKC]{Do you mean we should not use the word ``proof'' everywhere in this 
paper?}

>>>>

\NOTE[DPF]{``current'' was for BFSser if we need it, which we probably won't
since the re-emergence of BFSser.  So let's reserve the word just in case.}

@@

The miniKanren in TRS2 implements \textbf{i}nterleaving DFS (\DFSi), the cause of this 
unexpected result. With this search strategy, each \conde{} clause takes half 
of its received computational resources and passes the other half to its 
following clauses, except for the last clause that takes all resources it 
receives. In the example above, the \texttt{a} clause takes half of all 
resourses. And the \texttt{b} clause takes a quarter. Thus \texttt{c} and 
\texttt{d} barely contribute to the result.

@@

Given the following program, it is natural to expect lists of each
letter to constitute $1/4$ in the answer list. \DFSi, TRS2's search
strategy, however, results in many more lists of \texttt{a}s than
lists of other letters. And some letters (e.g. \texttt{c} and
\texttt{d}) are rarely seen. The more clauses, the worse the situation.

@@

This implementation can be improved in two ways. First, as mentioned in 
\autoref{sec:fairconj}, \BFS{} puts answers in buckets and answers of the same 
cost are in the same bucket. In the above implementation, however, 
it is not obvious how we manage cost information---the \texttt{car}s of a 
space have cost 0 (i.e., they are all in the same bucket), and every 
thunk indicates an increment in cost. It is even more subtle that 
\appendInfFair{} and the \appendMapInfFair{} respects the cost information. 
Second, \appendInfFair{} is extravagant in memory usage. It makes $O(n+m)$ new 
\texttt{cons} cells every time, where $n$ and $m$ are the sizes of 
the first buckets of two input spaces. \DFSf{} is also space extravagant.

@@

@@

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are semantically equivalent, since the result of \texttt{(run n ? g)} is 
the same either way. (See supplements for the formal proof.) To 
compare efficiency, we translate \BFSser{}'s Haskell code into Racket (See 
supplements for the translated code). The translation is direct due to the 
similarity of the two relational programming systems. The translated code is 
longer than \BFSimp{}. And it runs slower in all benchmarks. Details about 
differences in efficiency are in \autoref{sec:quan}.

@@

@@

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}.
Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
or better. 
It is hard to conclude how \disj{} fairness affects performance based on these 
statistics.
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

\NOTE[DPF]{Please leave ``prove'' (just here).}
